Optimal (Multiway) Spatial Joins

### 2024-05-27 msj.tex
yufei

The theorem implies a recursive nature of $k$-SJ. Indeed, we will see that an $k$-SJ instance with input size $n$ and output size $\out$ can be converted to $O(k^3)$ instances of the $(k-1)$-SJ problem --- all having input size at most $n$ and output size at most $\out$ --- plus an additional processing cost of $O(n \log n + \out)$. Moreover, if $\A$ is comparison-based, so is the $k$-SJ algorithm that ensues. For 2-SJ, we can set $\A$  to the ``folklore algorithm'' mentioned in Section~\ref{sec:intro:prev}, which ensures $F_2(n, \out) = O(n \log n + \out)$. Combining this with \eqref{eqn:main:reccurrence} gives a recurrence that relates the time complexity of $k$-SJ to that of $(k-1)$-SJ. Solving the recurrence yields:

@@

%As we will see, conquering this sepcial version is the most challenging step in settling the original $k$-SJ problem optimally.

@@

\begin{lemma} \label{lmm:hv}
    Equipped with the algorithm $\A$ described above, the H-V $k$-SJ problem can be solved in time
    \myeqn{
        O(k) \cdot \big( F_{k-1}(n, \out) + n\log n + k \cdot \out \big) \nn
    }
    where $n$ (resp., $\out$) is the input (resp., output) size of the problem.
\end{lemma}

@@

As mentioned in Section~\ref{sec:intro:prev}, 2-SJ can be solved in $F_2(n, \out) = O(n \log n + \out)$ time using a comparison-based algorithm. By Theorem~\ref{thm:main-recur}, in general, a comparison-based $(k-1)$-SJ algorithm with runtime $F_{k-1}(n, \out)$ spawns a comparison-based $k$-SJ algorithm whose running time $F_k(n, \out)$ obeys \eqref{eqn:main:reccurrence}. Specifically, for $k \ge 3$, there is a constant $c > 1$ such that
\myeqn{
    && F_{k}(n, \out) \nn \\
    &\le& c \cdot k^3 \cdot \big( F_{k-1}(n, \out) + n \log n + k \cdot \out \big) \nn \\
    &=& c^2 \cdot k^3(k-1)^3 \cdot (F_{k-2}(n,\out) + n\log n + k\cdot \out) + \nn \\
    && c \cdot k^3(n\log n + k\cdot \out ) \nn\\
    &<& c^2 \cdot k^3(k-1)^3 \cdot F_{k-2}(n,\out) +  \nn\\
    && (c+c^2)\cdot k^3(k-1)^3(n\log n + k \cdot\out) \nn \\
    &\leq& c^3 \cdot k^3(k-1)^3(k-2)^3 \cdot F_{k-3}(n,\out) + \nn\\
    && (c+c^2+c^3)\cdot k^3(k-1)^3(k-2)^3(n\log n + k \cdot\out) \nn\\
    &\leq& ... \nn \\
    &\leq& c^{k-2} \cdot (k!)^3 \cdot F_{2}(n,\out) + \nn \\
    && \Big(\sum_{i =1}^{k-2}c^i \Big)\cdot (k!)^3\cdot (n\log n + k \cdot\out) \nn \\
    &\leq & 2c^{k-1} \cdot (k!)^3 \cdot O(n\log n + k \cdot\out) \nn
}
which completes the proof of Theorem~\ref{thm:main-alg}.

>>>>

The theorem implies a recursive nature of $k$-SJ. Indeed, we will see that an $k$-SJ instance with input size $n$ and output size $\out$ can be converted to $O(k^3)$ instances of the $(k-1)$-SJ problem --- all having input size at most $n$ and output size at most $\out$ --- plus an additional processing cost of $O(n \log n + \out)$. For 2-SJ, we can set $\A$  to the ``folklore algorithm'' mentioned in Section~\ref{sec:intro:prev}, which ensures $F_2(n, \out) = O(n \log n + \out)$. Combining this with \eqref{eqn:main:reccurrence} gives a recurrence that relates the time complexity of $k$-SJ to that of $(k-1)$-SJ. Solving the recurrence yields:

@@

@@

\begin{lemma} \label{lmm:hv}
    Equipped with the algorithm $\A$ described above, the H-V $k$-SJ problem can be solved in time
    \myeqn{
        O(k) \cdot \big( F_{k-1}(n, \out) + n\log n + k \cdot \out \big) \nn
    }
    where $n$ (resp., $\out$) is the input (resp., output) size of the problem. Furthermore, if $\A$ is comparison-based, so is the H-V $k$-SJ algorithm.
\end{lemma}

@@

As mentioned in Section~\ref{sec:intro:prev}, 2-SJ can be solved in $F_2(n, \out) = O(n \log n + \out)$ time using a comparison-based algorithm. By Theorem~\ref{thm:main-recur}, in general, a comparison-based $(k-1)$-SJ algorithm with runtime $F_{k-1}(n, \out)$ spawns a comparison-based $k$-SJ algorithm whose running time $F_k(n, \out)$ obeys \eqref{eqn:main:reccurrence}. Specifically, for $k \ge 3$, there is a constant $c > 1$ such that
\myeqn{
    && F_{k}(n, \out) \nn \\
    &\le& c \cdot k^3 \cdot \big( F_{k-1}(n, \out) + n \log n + k \cdot \out \big) \nn \\
    &=& c^2 \cdot k^3(k-1)^3 \cdot (F_{k-2}(n,\out) + n\log n + k\cdot \out) + \nn \\
    && c \cdot k^3 \cdot (n\log n + k\cdot \out ) \nn\\
    &<& c^2 \cdot k^3(k-1)^3 \cdot F_{k-2}(n,\out) +  \nn\\
    && (c+c^2)\cdot k^3(k-1)^3 \cdot (n\log n + k \cdot\out) \nn \\
    &\leq& c^3 \cdot k^3(k-1)^3(k-2)^3 \cdot F_{k-3}(n,\out) + \nn\\
    && (c+c^2+c^3)\cdot k^3(k-1)^3(k-2)^3 \cdot (n\log n + k \cdot\out) \nn\\
    &\leq& ... \nn \\
    &\leq& c^{k-2} \cdot (k!)^3 \cdot F_{2}(n,\out) + \nn \\
    && \Big(\sum_{i =1}^{k-2}c^i \Big)\cdot (k!)^3 \cdot (n\log n + k \cdot\out) \nn \\
    &\leq & 2c^{k-1} \cdot (k!)^3 \cdot O(n\log n + k \cdot\out) \nn
}
which completes the proof of Theorem~\ref{thm:main-alg}.
